\LARGE{\bf{Natural variation in teosinte at the domestication locus \emph{teosinte branched1} (\emph{tb1})}}\footnotemark

### 2014-11-05 AJB_Res2.tex
Edits
Edits to manuscript and response notes

\lev{\bf{DONE}} Page 10, line 12: Why is culm diameter not believed to correlated with tillering index?  Is there a reference for this? What do your results show?  More importantly, if you want to prove that Hopscotch has more effect on tillering than on other independent traits you really need to do more comprehensive phenotyping.  Please explain more fully what this particular analysis is meant to achieve.
["Culm diameter is not believed to be correlated with tillering index or variation at tb1 and is used as our independent trait for phenotyping analyses."]
\lev{I think we'd cite studer for this. we wanted a trait that would be independent so that we could compare genotype + culm patterns to genotype + tillering patterns.}\\
Discussion
Page 14, 3rd para: This paragraph starts off by examining explanations of genetic drift and natural selection for different frequencies of Hopscotch alleles amongst teosinte populations.  After discussing evidence for bottlenecks, the argument shifts to the lack of diversity in the 5' UTR of the tb1 gene, concluding that the results are consistent with the action of selection on the 5'UTR, and that there might be an ecological role for the gene in some populations.  This is a classic example of the veering in focus between the purported object of this study, hopscotch elements, and the tb1 gene itself.  In the end there is no explanation given for the hopscotch frequencies, instead the focus is switched back to the gene itself.  To the reader there is an implication (perhaps unintentional) that there is a causative link between Hopscotch and variation in the tb1 alleles, whereas I would say this has simply not been shown in teosinte, and the evidence in this manuscript would
argue against such a link.
\lev{I don't mind this paragraph. It ties in findings from previous studies that identified evidence of bottleneck in some of these pops with our findings, which i think is useful. We state that a population bottleneck could explain the pattern that we see in SLO, but that it cannot account for this pattern across multiple populations in the Jalisco region.}

>>>>

\lev{\bf{DONE}} Page 10, line 12: Why is culm diameter not believed to correlated with tillering index?  Is there a reference for this? What do your results show?  More importantly, if you want to prove that Hopscotch has more effect on tillering than on other independent traits you really need to do more comprehensive phenotyping.  Please explain more fully what this particular analysis is meant to achieve.
["Culm diameter is not believed to be correlated with tillering index or variation at tb1 and is used as our independent trait for phenotyping analyses."]
\lev{This is cited as Briggs et al. 2007 in the text}\\
Discussion
Page 14, 3rd para: This paragraph starts off by examining explanations of genetic drift and natural selection for different frequencies of Hopscotch alleles amongst teosinte populations.  After discussing evidence for bottlenecks, the argument shifts to the lack of diversity in the 5' UTR of the tb1 gene, concluding that the results are consistent with the action of selection on the 5'UTR, and that there might be an ecological role for the gene in some populations.  This is a classic example of the veering in focus between the purported object of this study, hopscotch elements, and the tb1 gene itself.  In the end there is no explanation given for the hopscotch frequencies, instead the focus is switched back to the gene itself.  To the reader there is an implication (perhaps unintentional) that there is a causative link between Hopscotch and variation in the tb1 alleles, whereas I would say this has simply not been shown in teosinte, and the evidence in this manuscript would
argue against such a link.
\lev{I don't mind this paragraph. It ties in findings from previous studies that identified evidence of bottleneck in some of these pops with our findings, which i think is useful. We state that a population bottleneck could explain the pattern that we see in SLO, but that it cannot account for this pattern across multiple populations in the Jalisco region.}

### 2014-11-05 manuscript.tex
Edits
Edits to manuscript and response notes

\textbf{Phenotyping---}To assess the contribution of \emph{tb1} to phenotypic variation in tillering in a natural population, we grew plants from seed sampled from the San Lorenzo population of \emph{parviglumis}, which had a high mean frequency (0.44) of the \emph{Hopscotch} insertion based on our initial genotyping. 
We measured tiller number and tillering index, the ratio of the sum of tiller lengths to plant height, for 206 plants from within the San Lorenzo population, and genotyped plants for the \emph{Hopscotch} insertion. 
We also measured culm diameter, a phenotype that differs between maize and teosinte \citep{Briggs2007} but has not been shown to be affected by the \emph{Hopscotch} insertion. Culm diameter is meant to be an independent trait against which we can compare patterns of tillering index x \emph{Hopscotch} genotype data. If tillering index in \emph{parviglumis} is affected by the \emph{Hopscotch} insertion the expectation is that patterns of tillering index data will have a significant correlation with \emph{Hopscotch} genotype, whereas we should find no significant correlation between culm diameter and \emph{Hopscotch} genotype. 
Phenotypic data are available at \url{http://dx.doi.org/10.6084/m9.figshare.776926}. 
Our plantings produced 82 homozygotes for the \emph{Hopscotch} insertion at \emph{tb1}, 104 heterozygotes, and 20 homozygotes lacking the insertion; these numbers do not deviate from expectations of Hardy-Weinberg equilibrium. 
After performing a repeated measures ANOVA between our transformed tillering index data and \emph{Hopscotch} genotype we find no significant correlation between genotype at the \emph{Hopscotch} insertion and tillering index (Fig. \ref{Fig4Boxplots}), tiller number, or culm diameter. 
Only on day 40 did we observe a weak but statistically insignificant ($r^2= 0.02$, $p=0.0848$) correlation between tillering index and the \emph{Hopscotch} genotype, although in the opposite direction of that expected, with homozygotes for the insertion showing a higher tillering index.

>>>>

\textbf{Phenotyping---}To assess the contribution of \emph{tb1} to phenotypic variation in tillering in a natural population, we grew plants from seed sampled from the San Lorenzo population of \emph{parviglumis}, which had a high mean frequency (0.44) of the \emph{Hopscotch} insertion based on our initial genotyping. 
We measured tiller number and tillering index, the ratio of the sum of tiller lengths to plant height, for 206 plants from within the San Lorenzo population, and genotyped plants for the \emph{Hopscotch} insertion. 
We also measured culm diameter, a phenotype that differs between maize and teosinte but has not been shown to be affected by the \emph{Hopscotch} insertion \citep{Briggs2007}. Culm diameter is meant to be an independent trait against which we can compare patterns of tillering index x \emph{Hopscotch} genotype data. If tillering index in \emph{parviglumis} is affected by the \emph{Hopscotch} insertion the expectation is that patterns of tillering index data will have a significant correlation with \emph{Hopscotch} genotype, whereas we should find no significant correlation between culm diameter and \emph{Hopscotch} genotype. 
Phenotypic data are available at \url{http://dx.doi.org/10.6084/m9.figshare.776926}. 
Our plantings produced 82 homozygotes for the \emph{Hopscotch} insertion at \emph{tb1}, 104 heterozygotes, and 20 homozygotes lacking the insertion; these numbers do not deviate from expectations of Hardy-Weinberg equilibrium. 
After performing a repeated measures ANOVA between our transformed tillering index data and \emph{Hopscotch} genotype we find no significant correlation between genotype at the \emph{Hopscotch} insertion and tillering index (Fig. \ref{Fig4Boxplots}), tiller number, or culm diameter. 
Only on day 40 did we observe a weak but statistically insignificant ($r^2= 0.02$, $p=0.0848$) correlation between tillering index and the \emph{Hopscotch} genotype, although in the opposite direction of that expected, with homozygotes for the insertion showing a higher tillering index.
